---
title: "Analyzing Refinery Sector Product Flow of Light Liquid Fuels"
subtitle: "AIDRES, JRC-IDEES, EUROSTAT"
date: last-modified
date-format: long
author: 
 - name: Partha Das
   affiliation: 
    - name: 'VITO'
      url: https://vito.be/nl
 - name: Pieter Valkering
   affiliation: 
    - name: 'VITO'
      url: https://vito.be/nl
 - name: Hans Van de Put
   affiliation: 
    - name: 'VITO'
      url: https://vito.be/nl
format: 
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
    theme: journal
bibliography: references.bib
keywords:
  - AIDRES
  - EUROSTAT
  - JRC-IDEES
license: "CC BY"
citation: true
funding: 
  statement: "EU Project MOPO"
---

# Introduction

This is a stand-alone quarto document with code chunks and documentation for comparing Refinery sector product flow in AIDRES/AIDRES 2.0, EUROSTAT, and JRC-IDEES databases.

This document extracts refinery product flow for certain fuels (to match with Light Liquid Fuel category of AIDRES) from JRC-IDEES, EUROSTAT, and AIDRES/AIDRES 2.0 database and generate comparison plots and tables to visualize and compare the product flow across those databases.

# Using this file

The files are produced using [R](https://posit.co/download/rstudio-desktop/) and [RStudio](https://posit.co/download/rstudio-desktop/).
Though this document can be executed using command line, it is recommended to use `RStudio` for rendering.

Clone this repository to your computer. Open the `mopo_repo.Rproj` (`RStudio needs to be installed`) file. Open `refinery_analysis.qmd` and click on `render button` (Crtl + Shift + K).

# Loading libraries

Checking whether the required packages are installed or not. If not, install them and load them. If they are already installed, just load them.

```{r}
#| warning: false

mypackages <- c('tidyverse', 'readxl', 'janitor', 'archive', 'DBI', 'odbc',
              'RPostgres', 'eurostat', 'stringr', 'forcats', 'writexl', 'xlsx', 'gt', 'here')

for (p in mypackages){
  if (!require(p, character.only = TRUE))
    {
    install.packages(p)
    library(p, character.only = TRUE)
  }
  }
```

# Reading country names and codes

This is provided with the repository by default in the `input_data\europe_country_codes` directory by default.

```{r}
country_codes <- read_xlsx(here('input_data', 'europe_country_codes', 'europe_country_code.xlsx'), sheet = 1)

country_codes
```

# JRC-IDEES Refinary product flow

## Download IDEES excel files

`JRC‑IDEES provides a consistent set of disaggregated energy-economy-emissions data for each Member State of the European Union, covering all sectors of the energy system for the 2000-2021 period: industry, buildings, transport, and power generation. This data complies with Eurostat energy balances while providing a plausible decomposition of energy consumption into specific processes and end uses` [@jrc-idees_report]. 

Checking whether there is a directory (`./input_data/jrc_idees_2021/energy_balance`) containing JRC-IDEES excel files for energy balance. If not (only checking with `JRC-IDEES-2021_EnergyBalance_AT.xlsx`), creating the directory, downloading the zipped files, and extracting them for each country.

```{r}

idees_data <- here('input_data', 'jrc_idees_2021', 'energy_balance')

if (!dir.exists(idees_data)){
  
  print("Directory doesn't exists! Creating one")
  
  dir.create(idees_data, recursive = TRUE)
  
  print("Downloading files")
  
  for(country_code in country_codes$country_iso_code){
  
  zip_file <- paste0("https://jeodpp.jrc.ec.europa.eu/ftp/jrc-opendata/JRC-IDEES/JRC-IDEES-2021_v1/JRC-IDEES-2021_", country_code)
  filename <- paste0("JRC-IDEES-2021_EnergyBalance_", country_code, ".xlsx")

  archive_extract(archive = zip_file,
                  dir = idees_data,
                  files = filename)
  }
  
  } else {
    
    print(paste('Directory', idees_data,  'already exists!'))
  
  if(!file.exists(here(idees_data, 'JRC-IDEES-2021_EnergyBalance_AT.xlsx'))){
    
     print("But files does not exists! downloading them")
  
    for(country_code in country_codes$country_iso_code){
  
  zip_file <- paste0("https://jeodpp.jrc.ec.europa.eu/ftp/jrc-opendata/JRC-IDEES/JRC-IDEES-2021_v1/JRC-IDEES-2021_", country_code)
  filename <- paste0("JRC-IDEES-2021_EnergyBalance_", country_code, ".xlsx")

  archive_extract(archive = zip_file,
                  dir = idees_data,
                  files = filename)
  }
  } else {
    
    print(paste('Files also already exist inside the directory, not downloading them again!'))
    
    }
}
```

## Refinery product flow

From the Energy balance excel workbook we focus on the `TO_RPI_RO` sheet.

For the light liquid fuel category we only focus on the following fuels:

-   `Motor gasoline (excluding biofuel portion)`
-   `Aviation gasoline`
-   `Gasoline-type jet fuel`
-   `Kerosene-type jet fuel (excluding biofuel portion)`
-   `Other kerosene`
-   `Gas oil and diesel oil (excluding biofuel portion)`

```{r}
read_data_ref_to <- function(country){

  file_path <- paste0(idees_data, '/JRC-IDEES-2021_EnergyBalance_', country, '.xlsx')
  
  read_excel(file_path, sheet = 'TO_RPI_RO', range = 'A2:X135',
             col_names = c('fuel', 'code', 2000:2021)) |>
    select(1, 2, 21) |> 
    mutate(country_code = country) |> 
    slice(36, 37, 38, 39, 40, 42, ) |> 
    pivot_longer(cols = `2018`, names_to = 'year', values_to = 'ktoe') |> 
    mutate(belongs_to = 'TO_RPI_RO') |> 
    mutate(gj_idees = ktoe * 1000 * 41.87,
           ktoe = NULL) |>
    group_by(country_code, fuel, code, belongs_to) |>
    summarise(gj_idees=sum(gj_idees), .groups = 'drop')
}

test_list_ref_to <- lapply(country_codes$country_iso_code, read_data_ref_to)

ref_energy_to_idees <- bind_rows(test_list_ref_to, .id = "id") |> select(-id) |> 
  select(-code, -belongs_to) |> 
  pivot_wider(names_from = fuel, values_from =  gj_idees) |>
  clean_names() |> 
  rename('motor_gasoline' = motor_gasoline_excluding_biofuel_portion,
         'aviation_gasoline' = aviation_gasoline,
         'jet_gasoline' = gasoline_type_jet_fuel,
         'jet_kerosene' = kerosene_type_jet_fuel_excluding_biofuel_portion,
         'other_kerosene' = other_kerosene,
         'gas_oil_diesel_oil' = gas_oil_and_diesel_oil_excluding_biofuel_portion) |> 
  pivot_longer(col = aviation_gasoline:other_kerosene , names_to = 'fuel', values_to = 'values') |>
  mutate(pj_idees = values/ 10^6 , .keep = 'unused',
       year = 2018,
       fuel = 'light_liquid_fuel') |> 
  group_by(country_code, fuel, year) |> 
  reframe(pj_ref = sum(pj_idees)) |> 
  mutate(source = 'idees') |> 
  left_join(country_codes, by =join_by('country_code' == 'country_iso_code')) |> 
  select(country_code, country, year, source, everything())

ref_energy_to_idees
```

# EUROSTAT Refinery product flow

## Downloading Eurostat energy balance

EUROSTAT API is used to dynamically fetch EUROSTAT data using `eurostat` package [@RJ-2017-019]. Please refer this [link](https://ropengov.github.io/eurostat/index.html) to more about the usage of the `eurostat` package.

```{r}
#| warning: false
#| output: false

dat_aggregated_ref <- get_eurostat('nrg_bal_c', filters = list(time = 2018),
                                   time_format = "num", type = 'label', 
                                   stringsAsFactors = TRUE)
```

## Refinery product flow

We only focus on `Transformation output - refineries and petrochemical industry - refinery output` (`TO_RPI_RO`). Fuel choices (`siec`) are similar to JRC-IDEES.

```{r}

ref_output_eurostat <- dat_aggregated_ref |> 
  filter(unit == 'Terajoule') |> 
  filter(siec %in% c('Motor gasoline (excluding biofuel portion)', 'Aviation gasoline', 'Gasoline-type jet fuel', 
                     'Kerosene-type jet fuel (excluding biofuel portion)', 
                     'Other kerosene', 'Gas oil and diesel oil (excluding biofuel portion)')) |> 
  filter(nrg_bal %in% c('Transformation output - refineries and petrochemical industry - refinery output')) |> 
  pivot_wider(names_from = siec, values_from =  values) |> 
  clean_names() |> 
  rename('motor_gasoline' = motor_gasoline_excluding_biofuel_portion,
         'aviation_gasoline' = aviation_gasoline,
         'jet_gasoline' = gasoline_type_jet_fuel,
         'jet_kerosene' = kerosene_type_jet_fuel_excluding_biofuel_portion,
         'other_kerosene' = other_kerosene,
         'gas_oil_diesel_oil' = gas_oil_and_diesel_oil_excluding_biofuel_portion) |> 
  
  pivot_longer(col = aviation_gasoline:gas_oil_diesel_oil, names_to = 'fuel', values_to = 'values') |>
  pivot_wider(names_from = nrg_bal, values_from =  values) |> 
  clean_names() |> 
  rename("TO_RPI_RO" = transformation_output_refineries_and_petrochemical_industry_refinery_output) |>
  mutate(net = TO_RPI_RO / 1000, .keep = 'unused') |> 
  rename('country_code' = geo,
             'year' = time,
             'ref_eurostat_pj' =  net) |> 
  arrange(country_code)|> 
  group_by(country_code, year) |> 
  summarise(pj_ref = sum(ref_eurostat_pj), .groups = 'drop') |> 
  mutate(fuel = 'light_liquid_fuel') |> 
  left_join(country_codes, by =join_by('country_code' == 'country')) |> 
  filter(!is.na(country_iso_code)) |> 
  select(country_code, country_iso_code, fuel, year, pj_ref) |> 
  rename('country' = country_code,
          'country_code'  = country_iso_code) |> 
  mutate(source = 'eurostat') |> 
  select(country_code, country, fuel, year, source, pj_ref )
  
ref_output_eurostat
```

# AIDRES + AIDRES 2.0 refinery product flow

### Establishing AIDRES database connection and fetching tables

AIDRES database is open source [@aidres_report]. However the data is available as a raw `.sql` file which needs to be further used to setup a PostgreSQL database connection to fetch various tables. The data is also available as an Excel workbook. However, it is very customized to be used for specific analysis.

### Importing AIDRES data from local `.csv` files

However, for an external user setting up a PostgreSQL server is not straight forward. Therefore, AIDRES tables are provided with this repository in the `input_data\aidres_data` directory.

the AIDRES tables used for this analysis are first exported into `.csv` files and then they are used further in the analysis. Following code chunks create the directory (`input_data\aidres_data`) for storing the AIDRES files, and exports the AIDRES data files into `.csv` format. However, they are currently deactivated as we provide the files already with the repo.

```{r}

aidres_data <- here('input_data', 'aidres_data')

if (!dir.exists(aidres_data)){
  
  print("AIDRES input data folder doesn't exist. Request VITO to provide the data")
  dir.create(aidres_data)
  
} else {
  
  print("AIDRES input data directory already exists. Not creating a new one!")
  print("AIDRES data tables also exist!")
  
}
```

The AIDRES data tables  namely `aidres_scenarios.csv`, `aidres_sectors.csv`, `aidres_result.csv`, `aidres_configurations.csv`, `aidres_perton.csv` are made available in the `aidres_data` folder. The code chunk is currently deactivated.

Dumping AIDRES PSQL tables to `.csv` files (currently deactivated)

```{r}
#| eval: false

aidres_scenarios |> collect() |> write_csv(here(aidres_data, 'aidres_scenarios.csv'))

aidres_sectors |> collect() |>  write_csv(here(aidres_data, 'aidres_sectors.csv'))

aidres_result |> collect() |>  write_csv(here(aidres_data, 'aidres_result.csv'))

aidres_configurations|> collect() |> write_csv(here(aidres_data, 'aidres_configurations.csv')) 

aidres_perton|> collect() |> write_csv(here(aidres_data, 'aidres_perton.csv'))
```

### Reading the AIDRES and AIDRES 2.0 data from `.csv` files. 

```{r}

#| warning: false

aidres_scenarios <- read_csv(here(aidres_data, 'aidres_scenarios.csv'))
aidres_sectors <- read_csv(here(aidres_data, 'aidres_sectors.csv'))
aidres_result <- read_csv(here(aidres_data, 'aidres_result.csv'))
aidres_configurations <- read_csv(here(aidres_data, 'aidres_configurations.csv'))
aidres_perton <- read_csv(here(aidres_data, 'aidres_perton.csv'))


aidres_2_0_sectors <- read_csv2(here(aidres_data, 'aidres_2_0_sectors.csv'))
aidres_2_0_results <- read_delim(here(aidres_data, "aidres_2_0_results.csv"),
                                 delim = ";",
                                 locale = locale(decimal_mark = ".", grouping_mark = ","),
                                 guess_max = 6000)
aidres_2_0_configurations <- read_csv2(here(aidres_data, 'aidres_2_0_configurations.csv'))
aidres_2_0_perton <- read_delim(here(aidres_data, "aidres_2_0_perton.csv"),
                                 delim = ";",
                                 locale = locale(decimal_mark = ".", grouping_mark = ","),
                                 guess_max = 6000)
aidres_2_0_products <- read_csv2(here(aidres_data, 'aidres_2_0_products.csv'))
```

## Production all sectors in 2018

Product flow of different products in various sectors

```{r}
#| warning: false

aidres_result |> clean_names() |> 
  pivot_longer('product_flow_kt_y':'total_energy_vectors_pj_y', 
               names_to = 'indicators', values_to = 'value') |>
  mutate(value = round(value,3)) |> 
  filter(indicators == 'product_flow_kt_y') |> 
  filter(is_ref_scenario == 1,
         is_ref_route == 1) |> 
  select(-c(is_pareto_solution, is_direct_emission_solution, is_mix)) |> 
  left_join(aidres_sectors, by =join_by('aidres_sector_id' == 'id')) |> 
  select(-description,) |> 
  left_join(aidres_configurations, by =join_by('aidres_sector_id' == 'aidres_sector_id', 
                                               'configuration_id' == 'configuration_id')) |> 
  select( c('nuts3', 'sector_id', 'product_id'), everything(), -'aidres_sector_id', -'name', -starts_with('mix'),
          -abbreviation, -description, -starts_with("is"), -'value', -'indicators', 'indicators', 'value') |> 
  select(-c(configuration_id,	solution_id,	scenario_id,	route_id)) |> 
  rename('production_route_name' = route_name) |> 
  pivot_wider(names_from = 'indicators', values_from = 'value') |> 
  arrange(nuts3, sector_id, product_id, production_route_name, .by_group = TRUE) |> 
  separate_wider_position(nuts3, c(country_code = 2, 1, nuts_id =2)) |> 
  group_by(country_code, sector_id, product_id, production_route_name) |> 
  summarise(production_kt_y = sum(product_flow_kt_y)) |>  ungroup() ->  aidres_production_country_2018

aidres_production_country_2018
```

```{r}
#| warning: false
aidres_2_0_results |> clean_names() |> 
  pivot_longer('product_flow_kt_y':'total_energy_vectors_pj_y', 
               names_to = 'indicators', values_to = 'value') |>
  mutate(value = round(value,3)) |> 
  filter(indicators == 'product_flow_kt_y') |> 
  filter(scenario_id == 0,
         is_reference_route == 1) |> 
  select(-c(is_mix, is_reference_route)) |> 
  left_join(aidres_2_0_sectors, by =join_by('sector_id' == 'sector_id')) |>
  left_join(aidres_2_0_configurations, by =join_by('configuration_id' == 'configuration_id')) |>
  left_join(aidres_2_0_products, by =join_by('product_id' == 'product_id')) |>
  select(c('nuts3', 'sector', 'product', 'configuration', 'indicators', 'value')) |>
  rename('sector_id' = sector) |> 
  rename('product_id' = product) |> 
  mutate(sector_id = tolower(sector_id)) |>
  filter(product_id != 'fertiliser-nitric-acid') |> 
  filter(product_id != 'fertiliser-urea') |> 
  rename('production_route_name' = configuration) |>
  pivot_wider(names_from = 'indicators', values_from = 'value')|>
  separate_wider_position(nuts3, c(country_code = 2, 1, nuts_id =2)) |>
  group_by(country_code, sector_id, product_id, production_route_name) |> 
  summarise(production_kt_y = sum(product_flow_kt_y)) |>  ungroup() ->  aidres_2_0_production_country_2018

aidres_2_0_production_country_2018

```

## Refinery product flow

AIDRES/AIDRES 2.0 only considers light liquid fuels with lower heating value (LHV) of 42.87 MJ/kg. AIDRES/AIDRES 2.0 product flow is reported in Million tonne. We are using the LHV value to convert the product flow into PJ.

```{r}
aidres_production_country_2018 |> 
  filter(sector_id == 'refineries') |> 
  select(-sector_id, -production_route_name) |> 
  mutate(mt = round(production_kt_y/1000, 3), .keep = 'unused') |> 
  pivot_wider(names_from = product_id, values_from = mt) |>
  clean_names() |> 
  pivot_longer(cols = refineries_light_liquid_fuel, names_to = 'product_id', values_to = 'mt') |> 
  mutate(pj_ref = mt * 1000 * 10^6 * 42.87/ 10^9,
         fuel = 'light_liquid_fuel',
          source = 'aidres',
          year = 2018) |> 
  left_join(country_codes, by =join_by('country_code' == 'country_iso_code')) |> 
  select(country_code, country, fuel, year, source, pj_ref) -> aidres_production_country_ref

aidres_production_country_ref
```

```{r}
aidres_2_0_production_country_2018 |> 
  filter(sector_id == 'refineries') |> 
  select(-sector_id) |>
  mutate(mt = round(production_kt_y/1000, 3), .keep = 'unused') |> 
  pivot_wider(names_from = product_id, values_from = mt) |>
  clean_names() |> 
  pivot_longer(cols = refineries_light_liquid_fuel, names_to = 'product_id',values_to = 'mt') |> 
  mutate(pj_ref = mt * 1000 * 10^6 * 42.87/ 10^9,
         fuel = 'light_liquid_fuel',
          source = 'aidres',
          year = 2018) |>
  left_join(country_codes, by =join_by('country_code' == 'country_iso_code')) |>
  select(country_code, country, fuel, year, source, pj_ref) -> aidres_2_0_production_country_ref

aidres_2_0_production_country_ref
```

# Joining the data from three sources

Prepare the data from the three sources for comparison.

```{r}

bind_rows(aidres_production_country_ref, ref_energy_to_idees, ref_output_eurostat) |> 
  arrange(country_code) -> ref_data_sources

ref_data_sources
```

```{r}

bind_rows(aidres_2_0_production_country_ref, ref_energy_to_idees, ref_output_eurostat) |> 
  arrange(country_code) -> ref_data_sources_2_0

ref_data_sources_2_0
```

# Data comparison

## Creating output directory

```{r}
output_dir <- here('outputs', 'refinery')

if (!dir.exists(output_dir)){
  print("Output directory doesn't exist. Creating a new one!")
dir.create(output_dir)
} else {
    print("Output directory already exists. Not creating a new one!")
}
```

## Plot

```{r}

#| warning: false

  ref_data_sources |>
    ggplot(aes(pj_ref, fct_reorder(country_code, pj_ref, .fun=sum))) +
    geom_line()+
    geom_point(aes(colour = source), size = 1.5) +
    theme_bw() +
    labs(title = 'Light liquid fuel production in Refineries in AIDRES, EUROSTAT, and JRC-IDEES',
         y = 'Country',
         x = 'PJ') +
    theme(legend.position = "bottom") -> p1
    p1
    
    ggsave(here(output_dir, 'ref_pf_compare.pdf'), width = 14, height = 8, units = 'in', dpi = 900)
```

```{r}

#| warning: false

  ref_data_sources_2_0 |>
    ggplot(aes(pj_ref, fct_reorder(country_code, pj_ref, .fun=sum))) +
    geom_line()+
    geom_point(aes(colour = source), size = 1.5) +
    theme_bw() +
    labs(title = 'Light liquid fuel production in Refineries in AIDRES 2.0, EUROSTAT, and JRC-IDEES',
         y = 'Country',
         x = 'PJ') +
    theme(legend.position = "bottom") -> p1
    p1
    
    ggsave(here(output_dir, 'ref_pf_compare_2_0.pdf'), width = 14, height = 8, units = 'in', dpi = 900)
```

## Table

```{r}
ref_data_sources |> 
  select(-fuel, -year) |> 
  pivot_wider(names_from = source, values_from = pj_ref) |>
  mutate(across(.cols = -(1:2), .fns = ~replace_na(., replace = 0)),
         across(.cols = -(1:2), .fns = ~round(., 2))) |> 
  arrange(country_code) |>
  gt() |> 
  cols_label(
    country_code = 'Code',
    country = 'Name',
    aidres = 'AIDRES',
    idees = 'JRC-IDEES',
    eurostat = 'EUROSTAT'
  ) |> 
  tab_spanner(label = md('**Data Source**'),
    columns = 3:5) |> 
    tab_spanner(label = md('**Country**'),
    columns = 1:2) |> 
  tab_header(title = 'Light liquid fuel production (PJ) from refineries in 2018',
    subtitle = 'Comparison between AIDRES, JRC-IDEES, and EUROSTAT') |> 
  sub_zero(zero_text = '-') |> 
  opt_stylize(style = 6, color = 'gray')
```

```{r}
ref_data_sources_2_0 |> 
  select(-fuel, -year) |> 
  pivot_wider(names_from = source, values_from = pj_ref) |>
  mutate(across(.cols = -(1:2), .fns = ~replace_na(., replace = 0)),
         across(.cols = -(1:2), .fns = ~round(., 2))) |> 
  arrange(country_code) |>
  gt() |> 
  cols_label(
    country_code = 'Code',
    country = 'Name',
    aidres = 'AIDRES 2.0',
    idees = 'JRC-IDEES',
    eurostat = 'EUROSTAT'
  ) |> 
  tab_spanner(label = md('**Data Source**'),
    columns = 3:5) |> 
    tab_spanner(label = md('**Country**'),
    columns = 1:2) |> 
  tab_header(title = 'Light liquid fuel production (PJ) from refineries in 2018',
    subtitle = 'Comparison between AIDRES 2.0, JRC-IDEES, and EUROSTAT') |> 
  sub_zero(zero_text = '-') |> 
  opt_stylize(style = 6, color = 'gray')
```

### Diff file

The data containing product flow values from the three sources are exported into an Excel workbook for further analysis. `NA` values are replaced by `0`.

```{r}

ref_data_sources |> 
  select(-fuel, -year) |> 
  mutate(unit = 'pj') |> 
  select(country_code: source, unit, pj_ref) |> 
  pivot_wider(names_from = source, values_from = pj_ref) |>
  mutate(across(.cols = -(1:3), .fns = ~replace_na(., replace = 0)),
         across(.cols = -(1:3), .fns = ~round(., 1))) |> 
  arrange(country_code) |> 
   write_xlsx(here(output_dir,'diff_ref_prod_flow_llf.xlsx'))
```

```{r}

ref_data_sources_2_0 |> 
  select(-fuel, -year) |> 
  mutate(unit = 'pj') |> 
  select(country_code: source, unit, pj_ref) |> 
  pivot_wider(names_from = source, values_from = pj_ref) |>
  mutate(across(.cols = -(1:3), .fns = ~replace_na(., replace = 0)),
         across(.cols = -(1:3), .fns = ~round(., 1))) |> 
  arrange(country_code) |> 
   write_xlsx(here(output_dir,'diff_ref_prod_flow_llf_2_0.xlsx'))
```

# Bibliography