---
title: "Steel sector data analysis and input data preparation for EU/ Industrial Cluster case study in MOPO"
subtitle: "AIDRES, JRC-IDEES, EUROSTAT"
date: last-modified
date-format: long
author: 
 - name: Partha Das
   affiliation: 
    - name: 'VITO'
      url: https://vito.be/nl
 - name: Pieter Valkering
   affiliation: 
    - name: 'VITO'
      url: https://vito.be/nl
 - name: Hans Van de Put
   affiliation: 
    - name: 'VITO'
      url: https://vito.be/nl
format: 
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
    theme: journal
bibliography: references.bib
keywords:
  - AIDRES
  - EUROSTAT
  - JRC-IDEES
license: "CC BY"
citation: true
funding: 
  statement: "EU Project MOPO"
---

# Introduction

This is a stand-alone reproducible quarto document with code chunks and documentation for benchmarking Steel production and energy consumption in AIDRES, EUROSTAT.

It calibrates the AIDRES energy consumption w.r.t to EUROSTAT taking input from JRC-IDEES database. It generates comparison plots and tables to visualize and compare the steel production and energy consumption across those databases.

Finally, it calculates specific energy consumption values for additional processes (steel finishing, steel rolling) which are not part of AIDRES database. 

## Using this file

The files are produced using [R](https://posit.co/download/rstudio-desktop/) and [RStudio](https://posit.co/download/rstudio-desktop/).
Though this document can be executed using command line, it is recommended to use `RStudio` for rendering.

Clone this repository to your computer. Open the `mopo_repo.Rproj` (`RStudio needs to be installed`) file. Open `refinery_analysis.qmd` and click on `render button` (Crtl + Shift + K).

# Loading libraries

Checking whether the required packages are installed or not. If not, install them and load them. If they are already installed, just load them.

```{r}
#| warning: false

mypackages <- c('tidyverse', 'knitr', 'rmarkdown', 'readxl', 'janitor', 'archive', 'DBI', 'odbc', 'RPostgres', 'eurostat', 'stringr', 'forcats', 'writexl', 'xlsx', 'gt', 'here', 'svglite')

for (p in mypackages){
  if (!require(p, character.only = TRUE))
    {
    install.packages(p)
    library(p, character.only = TRUE)
  }
  }
```

# Reading country names and codes

This is provided with the repository by default in the `input_data\europe_country_codes` directory.

```{r}
country_codes <- read_xlsx(here('input_data', 'europe_country_codes', 'europe_country_code.xlsx'), sheet = 1)

country_codes
```

# AIDRES Steel production and energy consumption

AIDRES database is open source [@aidres_report]. However the data is available as a raw `.sql` file which needs to be further used to setup a PostgreSQL database connection to fetch various tables. The data is also available as an Excel workbook. However, it is very customized to be used for specific analysis.

## Importing AIDRES and AIDRES 2.0 data from local `.csv` files

However, for an external user setting up a PostgreSQL server is not straight forward. Therefore, AIDRES tables are provided with this repository in the `input_data\aidres_data` directory.

the AIDRES tables used for this analysis are first exported into `.csv` files and then they are used further in the analysis. Following code chunks create the directory (`input_data\aidres_data`) for storing the AIDRES files, and exports the AIDRES data files into `.csv` format. However, they are currently deactivated as we provide the files already with the repo.

```{r}

aidres_data <- here('input_data', 'aidres_data')

if (!dir.exists(aidres_data)){
  
  print("AIDRES input data folder doesn't exist. Request VITO to provide the data")
  dir.create(aidres_data)
  
} else {
  
  print("AIDRES input data directory already exists. Not creating a new one!")
  print("AIDRES data tables also exist!")
  
}
```

The AIDRES data tables namely `aidres_scenarios.csv`, `aidres_sectors.csv`, `aidres_result.csv`, `aidres_configurations.csv`, `aidres_perton.csv` are made available in the `input_data\aidres_data` folder. The code chunk is currently deactivated.

Dumping AIDRES PSQL tables to `.csv` files (One time run. Currently deactivated)

```{r}
#| eval: false

aidres_scenarios |> collect() |> write_csv(here(aidres_data, 'aidres_scenarios.csv'))

aidres_sectors |> collect() |>  write_csv(here(aidres_data, 'aidres_sectors.csv'))

aidres_result |> collect() |>  write_csv(here(aidres_data, 'aidres_result.csv'))

aidres_configurations|> collect() |> write_csv(here(aidres_data, 'aidres_configurations.csv')) 

aidres_perton|> collect() |> write_csv(here(aidres_data, 'aidres_perton.csv'))
```

### Reading the AIDRES data `.csv` files. 

```{r}
#| warning: false

aidres_scenarios <- read_csv(here(aidres_data, 'aidres_scenarios.csv'))
aidres_sectors <- read_csv(here(aidres_data, 'aidres_sectors.csv'))
aidres_result <- read_csv(here(aidres_data, 'aidres_result.csv'))
aidres_configurations <- read_csv(here(aidres_data, 'aidres_configurations.csv'))
aidres_perton <- read_csv(here(aidres_data, 'aidres_perton.csv'))

aidres_2_0_sectors <- read_csv2(here(aidres_data, 'aidres_2_0_sectors.csv'))
aidres_2_0_results <- read_delim(here(aidres_data, "aidres_2_0_results.csv"),
                                 delim = ";",
                                 locale = locale(decimal_mark = ".", grouping_mark = ","),
                                 guess_max = 6000)
aidres_2_0_configurations <- read_csv2(here(aidres_data, 'aidres_2_0_configurations.csv'))
aidres_2_0_products <- read_csv2(here(aidres_data, 'aidres_2_0_products.csv'))
```

## Steel production in 2018

First we extract the Production volumes of different products in various sectors from AIDRES database.

```{r}
#| warning: false

aidres_result |> clean_names() |> 
  pivot_longer('product_flow_kt_y':'total_energy_vectors_pj_y', names_to = 'indicators', values_to = 'value') |>
  mutate(value = round(value,3)) |> 
  filter(indicators == 'product_flow_kt_y') |> 
  filter(is_ref_scenario == 1,
         is_ref_route == 1) |> 
  select(-c(is_pareto_solution, is_direct_emission_solution, is_mix)) |> 
  left_join(aidres_sectors, by =join_by('aidres_sector_id' == 'id')) |> 
  select(-description,) |> 
  left_join(aidres_configurations, by =join_by('aidres_sector_id' == 'aidres_sector_id', 
                                               'configuration_id' == 'configuration_id')) |> 
  select( c('nuts3', 'sector_id', 'product_id'), everything(), -'aidres_sector_id', -'name', -starts_with('mix'),
          -abbreviation, -description, -starts_with("is"), -'value', -'indicators', 'indicators', 'value') |> 
  select(-c(configuration_id,	solution_id,	scenario_id,	route_id)) |> 
  rename('production_route_name' = route_name) |> 
  pivot_wider(names_from = 'indicators', values_from = 'value') |> 
  arrange(nuts3, sector_id, product_id, production_route_name, .by_group = TRUE) |> 
  separate_wider_position(nuts3, c(country_code = 2, 1, nuts_id =2)) |> 
  group_by(country_code, sector_id, product_id, production_route_name) |> 
  summarise(production_kt_y = sum(product_flow_kt_y)) |>  ungroup() -> aidres_production_country_2018

  aidres_production_country_2018
```

First we extract the Production volumes of different products in various sectors from AIDRES 2.0 database.

```{r}
#| warning: false
aidres_2_0_results |> clean_names() |> 
  pivot_longer('product_flow_kt_y':'total_energy_vectors_pj_y', 
               names_to = 'indicators', values_to = 'value') |>
  mutate(value = round(value,3)) |> 
  filter(indicators == 'product_flow_kt_y') |> 
  filter(scenario_id == 0,
         is_reference_route == 1) |> 
  select(-c(is_mix, is_reference_route)) |> 
  left_join(aidres_2_0_sectors, by =join_by('sector_id' == 'sector_id')) |>
  left_join(aidres_2_0_configurations, by =join_by('configuration_id' == 'configuration_id')) |>
  left_join(aidres_2_0_products, by =join_by('product_id' == 'product_id')) |>
  select(c('nuts3', 'sector', 'product', 'configuration', 'indicators', 'value')) |>
  rename('sector_id' = sector) |> 
  rename('product_id' = product) |> 
  mutate(sector_id = tolower(sector_id)) |>
  filter(product_id != 'fertiliser-nitric-acid') |> 
  filter(product_id != 'fertiliser-urea') |> 
  rename('production_route_name' = configuration) |>
  pivot_wider(names_from = 'indicators', values_from = 'value')|>
  separate_wider_position(nuts3, c(country_code = 2, 1, nuts_id =2)) |>
  group_by(country_code, sector_id, product_id, production_route_name) |> 
  summarise(production_kt_y = sum(product_flow_kt_y)) |>  ungroup() ->  aidres_2_0_production_country_2018

aidres_2_0_production_country_2018
```

Filtering only Steel sector production data (AIDRES)

```{r}

isi_aidres_production_comp <- aidres_production_country_2018 |> 
  filter(sector_id == 'steel') |> 
  select(-sector_id) |> 
  pivot_longer(production_kt_y, values_to = 'kt', names_to = 'source') |> 
  mutate(source = 'aidres')  |> 
  pivot_wider(names_from = product_id, values_from = kt) |> 
  clean_names() |> 
  pivot_longer(cols = steel_primary: steel_secondary, names_to = 'product_id',values_to = 'kt') |> 
  filter(!is.na(kt))

        # mutate(steel_total = rowSums(cbind(steel_primary, steel_secondary), na.rm = TRUE)) 
        # pivot_longer(cols = steel_primary:steel_total, names_to = 'product_id', values_to = 'kt') |> 
        # mutate(kt = replace_na(kt, 0))
      
  isi_aidres_production_comp
```

Filtering only Steel sector production data (AIDRES 2.0)

```{r}

isi_aidres_2_0_production_comp <- aidres_2_0_production_country_2018 |> 
  filter(sector_id == 'steel') |> 
  select(-sector_id) |> 
  pivot_longer(production_kt_y, values_to = 'kt', names_to = 'source') |> 
  mutate(source = 'aidres')  |> 
  pivot_wider(names_from = product_id, values_from = kt) |> 
  clean_names() |> 
  pivot_longer(cols = steel_primary: steel_secondary, names_to = 'product_id',values_to = 'kt') |> 
  filter(!is.na(kt))

        # mutate(steel_total = rowSums(cbind(steel_primary, steel_secondary), na.rm = TRUE)) 
        # pivot_longer(cols = steel_primary:steel_total, names_to = 'product_id', values_to = 'kt') |> 
        # mutate(kt = replace_na(kt, 0))
      
  isi_aidres_2_0_production_comp
```

## Energy consumption for Steel production in 2018

First we calculate Energy consumption by different production routes in different sectors (AIDRES). 

```{r}
#| warning: false

aidres_result |> clean_names() |> 
  select(nuts3, contains('id'), contains('is_'), contains('_pj_y'))|>
  pivot_longer('electricity_pj_y':'total_energy_vectors_pj_y', names_to = 'indicators', values_to = 'value') |> 
  mutate(value = round(value,3)) |>
  filter(is_ref_scenario == 1,
         is_ref_route == 1) |> 
  select(-c(is_pareto_solution, is_direct_emission_solution, is_mix)) |> 
  left_join(aidres_sectors, by =join_by('aidres_sector_id' == 'id')) |> 
  select(-description) |> 
  left_join(aidres_configurations, by =join_by('aidres_sector_id' == 'aidres_sector_id', 
                                               'configuration_id' == 'configuration_id')) |> 
  select( c('nuts3', 'sector_id', 'product_id'), everything(), -'aidres_sector_id', -'name', -starts_with('mix'), -abbreviation, -description,
          -starts_with("is"), -'value', -'indicators', 'indicators', 'value') |> 
  select(-c(configuration_id,	solution_id,	scenario_id,	route_id, id)) |> 
  rename('production_route_name' = route_name) |> 
  select(nuts3, sector_id, product_id, production_route_name, everything()) |> 
  arrange(nuts3, sector_id, product_id, production_route_name, .by_group = TRUE) |> 
  separate_wider_position(nuts3, c(country_code = 2, 1, nuts_id =2))|>
  group_by(country_code, sector_id, indicators) |> 
  summarise(country_total = sum(value)) |> 
  ungroup() |> 
  left_join(country_codes, by =join_by('country_code' == 'country_iso_code')) |> 
  select(country_code, country, everything()) |> 
  pivot_wider(names_from = sector_id, values_from = country_total) -> aidres_energy_all

  aidres_energy_all
```

First we calculate Energy consumption by different production routes in different sectors (AIDRES 2.0). 

```{r}
#| warning: false

aidres_2_0_results |> clean_names() |> 
  pivot_longer('alternative_fuel_mixture_usage_pj_y':'total_energy_vectors_pj_y', 
               names_to = 'indicators', values_to = 'value') |>
  mutate(value = round(value,3)) |> 
  filter(scenario_id == 0,
         is_reference_route == 1) |> 
  select(-c(is_mix, is_reference_route)) |> 
  left_join(aidres_2_0_sectors, by =join_by('sector_id' == 'sector_id')) |>
  left_join(aidres_2_0_configurations, by =join_by('configuration_id' == 'configuration_id')) |>
  left_join(aidres_2_0_products, by =join_by('product_id' == 'product_id')) |>
  select(c('nuts3', 'sector', 'product', 'configuration', 'indicators', 'value')) |>
  rename('sector_id' = sector) |> 
  rename('product_id' = product) |> 
  mutate(sector_id = tolower(sector_id)) |>
  filter(product_id != 'fertiliser-nitric-acid') |> 
  filter(product_id != 'fertiliser-urea') |> 
  rename('production_route_name' = configuration) |>
  separate_wider_position(nuts3, c(country_code = 2, 1, nuts_id =2)) |>
  group_by(country_code, sector_id, indicators) |> 
  summarise(country_total = sum(value)) |>
  left_join(country_codes, by =join_by('country_code' == 'country_iso_code')) |> 
  pivot_wider(names_from = sector_id, values_from = country_total) -> aidres_2_0_energy_all

  aidres_2_0_energy_all
```

Filtering out Steel sector energy consumption and aggregating fuels to EUROSTAT aggregated commodities (AIDRES).

```{r}

aidres_cleaned_steel <- aidres_energy_all |> 
  select(country_code:indicators, steel) |> 
  mutate(indicators = str_replace(indicators, "\\_pj_y", "")) |> 
  rename('steel_aidres' = steel,
         'fuel' = indicators) |> 
  pivot_longer(cols = steel_aidres, names_to = 'metadata', values_to = 'pj') |> 
  pivot_wider(names_from = fuel, values_from = pj) |> 
  rename('total' = total_energy_vectors) |> 
  #aggregate to broad fuel categories of eurostat  
  mutate(solid_fossil_fuels = rowSums(cbind(coal, coke), na.rm = TRUE),
         renew_bio = rowSums(cbind(biomass, biomass_waste), na.rm = TRUE),
         non_renew_waste = rowSums(cbind(plastic_mix, alternative_fuel_mixture), na.rm = TRUE),
         oil_petro_products = rowSums(cbind(crude_oil, naphtha), na.rm = TRUE)) |> 
  select(-c(alternative_fuel_mixture, coal, coke, biomass, biomass_waste, plastic_mix, crude_oil, naphtha)) |> 
  pivot_longer(cols = -c(country_code:metadata), names_to = 'fuel', values_to = 'pj') |> 
  pivot_wider(names_from = metadata, values_from = pj)

aidres_cleaned_steel
```

Filtering out Steel sector energy consumption and aggregating fuels to EUROSTAT aggregated commodities (AIDRES 2.0).

```{r}

aidres_2_0_cleaned_steel <- aidres_2_0_energy_all |> 
  select(country_code:country, steel) |> 
  mutate(indicators = str_replace(indicators, "\\_usage_pj_y", "")) |> 
  mutate(indicators = str_replace(indicators, "\\_pj_y", "")) |> 
  rename('steel_aidres' = steel,
         'fuel' = indicators) |> 
  pivot_longer(cols = c(steel_aidres), names_to = 'metadata', values_to = 'pj') |> 
  pivot_wider(names_from = fuel, values_from = pj) |> 
  rename('total' = total_energy_vectors) |> 
  mutate(solid_fossil_fuels = rowSums(cbind(coal, coke, coal_steel), na.rm = TRUE),
         renew_bio = rowSums(cbind(biomass, biomass_waste), na.rm = TRUE),
         non_renew_waste = rowSums(cbind(plastic_mix, alternative_fuel_mixture), na.rm = TRUE),
         oil_petro_products = rowSums(cbind(crude_oil, naphtha), na.rm = TRUE)) |> 
  select(-c(alternative_fuel_mixture, coal, coke, biomass, biomass_waste, plastic_mix, crude_oil, naphtha, coal_steel)) |> 
  pivot_longer(cols = -c(country_code:metadata), names_to = 'fuel', values_to = 'pj') |>
  pivot_wider(names_from = metadata, values_from = pj)

  aidres_2_0_cleaned_steel
```

# EUROSTAT Steel energy consumption

## Downloading Eurostat energy balance

EUROSTAT API is used to dynamically fetch EUROSTAT data using `eurostat` package [@RJ-2017-019]. Please refer this [link](https://ropengov.github.io/eurostat/index.html) to more about the usage of the `eurostat` package.

For steel sector aggregated energy balance data of EUROSTAT (nrg_bal_c) is used.

```{r}
#| warning: false

dat_aggregated <- get_eurostat('nrg_bal_c', filters = list(time = 2018),
                                   time_format = "num", type = 'label', 
                                   stringsAsFactors = TRUE)
```

## Steel energy consumption

To calculate net energy consumption in the steel sector we consider the Transformation output, Transformation input, and Energy use for the coke ovens and blast furnace also.

```{r}
steel_energy_eurostat <- dat_aggregated |> 
  filter(unit == 'Terajoule') |> 
  filter(siec %in% c('Solid fossil fuels', 'Manufactured gases', 'Oil and petroleum products (excluding biofuel portion)', 'Renewables and biofuels', 'Electricity', 'Natural gas', 'Non-renewable waste', 'Heat', 'Total')) |> 
  pivot_wider(names_from = siec, values_from =  values) |> 
  clean_names() |> 
  rename('oil_petro_products' = oil_and_petroleum_products_excluding_biofuel_portion,
         'renew_bio' = renewables_and_biofuels,
         'non_renew_waste' = non_renewable_waste) |> 
  mutate(natural_gas  = rowSums(cbind(natural_gas , manufactured_gases), na.rm = TRUE)) |> 
  select(-manufactured_gases) |> 
  filter(nrg_bal %in% c('Final consumption - industry sector - iron and steel - energy use',
                         'Energy sector - coke ovens - energy use',
                         'Transformation input - coke ovens - energy use',
                         'Transformation output - coke ovens',
                         'Final consumption - industry sector - non-energy use',
                         'Energy sector - blast furnaces - energy use',
                         'Transformation input - blast furnaces - energy use',
                         'Transformation output - blast furnaces')) |> 
  pivot_longer(col = total:heat, names_to = 'fuel', values_to =  'values') |> 
  pivot_wider(names_from = nrg_bal, values_from =  values) |> 
  clean_names() |>
  rename("FC_IND_IS_E" = final_consumption_industry_sector_iron_and_steel_energy_use,
         "NRG_CO_E" = energy_sector_coke_ovens_energy_use,
         "NRG_BF_E" = energy_sector_blast_furnaces_energy_use,
         "TI_BF_E" = transformation_input_blast_furnaces_energy_use,
         "TI_CO_E" = transformation_input_coke_ovens_energy_use,
         "TO_CO" = transformation_output_coke_ovens,
         "TO_BF" = transformation_output_blast_furnaces,
         "FC_IND_NE" = final_consumption_industry_sector_non_energy_use) |> 
  group_by(fuel, geo, time) |> 
  mutate(net = (rowSums(cbind(FC_IND_IS_E, NRG_CO_E, NRG_BF_E, TI_BF_E, TI_CO_E), na.rm = TRUE) - rowSums(cbind(TO_CO, TO_BF), na.rm = TRUE))) |>
  mutate(net = net/1000) |> 
  rename('country_code' = geo,
         'year' = time,
         'steel_eurostat_pj' =  net) |> 
  arrange(country_code)|> 
  select(country_code, year, fuel, steel_eurostat_pj) |> 
  ungroup() |> 
  left_join(country_codes, by = join_by(country_code == country)) |> 
  filter(!is.na(country_iso_code)) |> 
  rename('country' = country_code,
         'country_code' = country_iso_code,
         'steel_eurostat' = steel_eurostat_pj) |> 
  select(country_code, country, fuel, everything(), -year) |> 
  arrange(country_code, fuel)

steel_energy_eurostat
```

## Comparison of Steel sector energy consumption between AIDRES/AIDRES 2.0 and EUROSTAT

AIDRES and AIDRES 2.0 data still do not have energy consumption from steel finishing and rolling.

### Creating output directory

The output directory `outputs\steel` is created to store all output files produced with this `.qmd` file. 

```{r}
output_dir <- here('outputs', 'steel')

if (!dir.exists(output_dir)){
  print("Output directory doesn't exist. Creating a new one!")
dir.create(output_dir)
} else {
  print("Output directory already exists. Not creating a new one!")
}
```

### Joining AIDRES/AIDRES 2.0 and EUROSTAT Steel energy consumption

```{r}

steel_eurostat_aidres <- steel_energy_eurostat |> 
  full_join(aidres_cleaned_steel, by = join_by(country_code, country, fuel)) |> 
  select(country_code, country, fuel, everything()) |> 
  pivot_longer(cols = steel_eurostat:steel_aidres, names_to = 'metadata', values_to = 'pj') |> 
  mutate(pj = round(pj, 2))|> 
  separate_wider_delim(metadata, delim = '_', names = c('sector', 'source'))

steel_eurostat_aidres 
```

```{r}

steel_eurostat_aidres_2_0 <- steel_energy_eurostat |> 
  full_join(aidres_2_0_cleaned_steel, by = join_by(country_code, country, fuel)) |> 
  select(country_code, country, fuel, everything()) |> 
  pivot_longer(cols = steel_eurostat:steel_aidres, names_to = 'metadata', values_to = 'pj') |> 
  mutate(pj = round(pj, 2))|> 
  separate_wider_delim(metadata, delim = '_', names = c('sector', 'source'))

steel_eurostat_aidres_2_0 
```

### Plot

Total energy consumption only

```{r,  fig.width=10,fig.height=8}
#| warning: false

  steel_eurostat_aidres |> filter(fuel == 'total') |>
    filter(sector == 'steel') |> 
    ggplot(aes(pj, fct_reorder(country_code, pj, .fun=sum))) +
    geom_line()+
    geom_point(aes(colour = source), size = 1.5) +
    theme_bw() +
    labs(title = 'Total energy consumption in the steel sector in AIDRES and EUROSTAT',
         subtitle = 'Rolling and finishing energy is missing in AIDRES database',
         y = 'country',
         x = 'PJ') +
    theme(legend.position = "bottom") +
    theme(legend.title = element_blank()) -> p3
    p3
    
     ggsave(here(output_dir, 'steel_fec_compare_tot_no_fr.pdf'), width = 14, height = 8, units = 'in', dpi = 900)
     
          ggsave(here(output_dir, 'steel_fec_compare_tot_no_fr.svg'), width = 12, height = 7, units = 'in', dpi = 900)
```

```{r,  fig.width=10,fig.height=8}
#| warning: false

  steel_eurostat_aidres_2_0 |> filter(fuel == 'total') |>
    filter(sector == 'steel') |> 
    ggplot(aes(pj, fct_reorder(country_code, pj, .fun=sum))) +
    geom_line()+
    geom_point(aes(colour = source), size = 1.5) +
    theme_bw() +
    labs(title = 'Total energy consumption in the steel sector in AIDRES 2.0 and EUROSTAT',
         subtitle = 'Rolling and finishing energy is missing in AIDRES 2.0 database',
         y = 'country',
         x = 'PJ') +
    theme(legend.position = "bottom") +
    theme(legend.title = element_blank()) -> p3
    p3
    
     ggsave(here(output_dir, 'steel_fec_compare_tot_no_fr_2_0.pdf'), width = 14, height = 8, units = 'in', dpi = 900)
     
          ggsave(here(output_dir, 'steel_fec_compare_tot_no_fr_2_0.svg'), width = 12, height = 7, units = 'in', dpi = 900)
```

Total energy consumption by fuel

```{r,  fig.width=10,fig.height=12}
#| warning: false

  steel_eurostat_aidres |>
    filter(sector == 'steel') |> 
    ggplot(aes(pj, fct_reorder(country_code, pj, .fun=sum))) +
    geom_line()+
    geom_point(aes(colour = source), size = 1.5) +
    facet_wrap(~fuel, ncol = 5, scales = "free") +
    theme_bw() +
    labs(title = 'Total energy consumption by fuel in the steel sector in AIDRES and EUROSTAT',
         subtitle = 'Rolling and finishing energy is missing in AIDRES database',
         y = 'country',
         x = 'PJ')+
    theme(legend.title = element_blank()) -> p20
    p20
    
     ggsave(here(output_dir, 'steel_fec_compare_fuel_no_fr.pdf'), width = 16, height = 8, units = 'in', dpi = 900)
     
          ggsave(here(output_dir, 'steel_fec_compare_fuel_no_fr.svg'), width = 12, height = 7, units = 'in', dpi = 900)
```

```{r,  fig.width=10,fig.height=12}
#| warning: false

  steel_eurostat_aidres_2_0 |>
    filter(sector == 'steel') |> 
    ggplot(aes(pj, fct_reorder(country_code, pj, .fun=sum))) +
    geom_line()+
    geom_point(aes(colour = source), size = 1.5) +
    facet_wrap(~fuel, ncol = 5, scales = "free") +
    theme_bw() +
    labs(title = 'Total energy consumption by fuel in the steel sector in AIDRES 2.0 and EUROSTAT',
         subtitle = 'Rolling and finishing energy is missing in AIDRES 2.0 database',
         y = 'country',
         x = 'PJ')+
    theme(legend.title = element_blank()) -> p20
    p20
    
     ggsave(here(output_dir, 'steel_fec_compare_fuel_no_fr_2_0.pdf'), width = 16, height = 8, units = 'in', dpi = 900)
     
          ggsave(here(output_dir, 'steel_fec_compare_fuel_no_fr_2_0.svg'), width = 12, height = 7, units = 'in', dpi = 900)
```

### Tables

Steel sector energy consumption by fuel

```{r}
steel_eurostat_aidres |> 
  select(-sector) |> 
  arrange(country_code, source, fuel) |> 
  pivot_wider(names_from = fuel, values_from = pj) |> 
  mutate(across(.cols = -(1:4), .fns = ~replace_na(., replace = 0)),
         across(.cols = -(1:4), .fns = ~round(., 2))) |> 
  arrange(country_code) |> 
  pivot_longer(electricity:total, names_to = 'fuel', values_to = 'pj') |> 
  pivot_wider(names_from = source, values_from = pj) -> steel_energy_eurostat_aidres

steel_energy_eurostat_aidres
```

```{r}
steel_eurostat_aidres_2_0 |> 
  select(-sector) |> 
  arrange(country_code, source, fuel) |> 
  pivot_wider(names_from = fuel, values_from = pj) |> 
  mutate(across(.cols = -(1:4), .fns = ~replace_na(., replace = 0)),
         across(.cols = -(1:4), .fns = ~round(., 2))) |> 
  arrange(country_code) |> 
  pivot_longer(electricity:total, names_to = 'fuel', values_to = 'pj') |> 
  pivot_wider(names_from = source, values_from = pj) -> steel_energy_eurostat_aidres_2_0

steel_energy_eurostat_aidres_2_0
```


### Diff files

The data containing product flow values from the three sources are exported into an Excel workbook for further analysis. `NA` values are replaced by `0`.

```{r}
steel_energy_eurostat_aidres |> 
   write_xlsx(here(output_dir, 'diff_no_finishing_rolling_fuel.xlsx'))
```

```{r}
steel_energy_eurostat_aidres_2_0 |> 
   write_xlsx(here(output_dir, 'diff_no_finishing_rolling_fuel_2_0.xlsx'))
```

# Adding rolling and finishing energy from JRC-IDEES

AIDRES/AIDRES 2.0 data still does not have energy consumption from steel finishing and rolling. JRC IDEES has breakup of total energy consumption of the whole steel sector in various sub-parts like rolling and finishing. Rolling and finishing data is taken from IRC IDEES and added on top of AIDRES final energy consumption by fuel. The resulting 'total' final energy is then compared with EUROSTAT data. Steel production data is also compared between JRC IDEES and AIDRES


## JRC-IDEES Steel Production volume, Rolling, and Finishing Energy

JRC-IDEES database is used first to produce historical trends of primary and secondary steel production in countries and then compare with AIDRES for the year 2018.

Subsequently, energy consumption for rolling and finishing is taken from JRC-IDEES and added to AIDRES/AIDRES 2.0 for calibration.

## Download IDEES excel files

`JRC‑IDEES provides a consistent set of disaggregated energy-economy-emissions data for each Member State of the European Union, covering all sectors of the energy system for the 2000-2021 period: industry, buildings, transport, and power generation. This data complies with Eurostat energy balances while providing a plausible decomposition of energy consumption into specific processes and end uses` [@jrc-idees_report]. 

Checking whether there is a directory (`./input_data/jrc_idees_2021/energy_balance`) containing JRC-IDEES excel files for energy balance. If not (only checking with `JRC-IDEES-2021_EnergyBalance_AT.xlsx`), creating the directory, downloading the zipped files, and extracting them for each country.

```{r}

idees_data <- here('input_data', 'jrc_idees_2021', 'industry')

if (!dir.exists(idees_data)){
  
  print("Directory doesn't exists! Creating one")
  
  dir.create(idees_data, recursive = TRUE)
  
  print("Downloading files")
  
  for(country_code in country_codes$country_iso_code){
  
  zip_file <- paste0("https://jeodpp.jrc.ec.europa.eu/ftp/jrc-opendata/JRC-IDEES/JRC-IDEES-2021_v1/JRC-IDEES-2021_", country_code)
  filename <- paste0("JRC-IDEES-2021_Industry_", country_code, ".xlsx")

  archive_extract(archive = zip_file,
                  dir = idees_data,
                  files = filename)
  }
  
  } else {
    
    print(paste('Directory', idees_data,  'already exists!'))
  
  if(!file.exists(here(idees_data, 'JRC-IDEES-2021_Industry_AT.xlsx'))){
    
     print("But files does not exists! downloading them")
  
    for(country_code in country_codes$country_iso_code){
  
  zip_file <- paste0("https://jeodpp.jrc.ec.europa.eu/ftp/jrc-opendata/JRC-IDEES/JRC-IDEES-2021_v1/JRC-IDEES-2021_", country_code)
  filename <- paste0("JRC-IDEES-2021_Industry_", country_code, ".xlsx")

  archive_extract(archive = zip_file,
                  dir = idees_data,
                  files = filename)
  }
  } else {
    
    print(paste('Files also already exist inside the directory, not downloading them again!'))
    
    }
}
```

## Primary and secondary steel production in 2018

Excel files have been copied manually into a folder inside the working directory. Creating functions to read all the excel files into a dataframe from a directory.

```{r}

read_data_isi_2018 <- function(country){
  
 file_path <- paste0(idees_data, '/JRC-IDEES-2021_Industry_', country, ".xlsx")
 
  read_excel(file_path, sheet = 'ISI', range = 'A2:W20',
             col_names = c('product', 2000:2021)) |>
  select(1,20) |> 
  mutate(country_code = country) |> 
  slice(7:8) |> 
  mutate(product = replace(product,
                           1:2, c('steel-primary', 'steel-secondary'))) |> 
  pivot_longer(cols ='2018', names_to = 'year', values_to = 'production_kt_idees')
}

test_list_isi_2018 <- lapply(country_codes$country_iso_code, read_data_isi_2018)
isi_production_2018 <- bind_rows(test_list_isi_2018, .id = "id") |> select(-id, -year)
```

### Yearly Total Steel production of countries (Historical years)

From JRC-IDEES

```{r}

read_data_isi_allyear<- function(country){
  
file_path <- paste0(idees_data, '/JRC-IDEES-2021_Industry_', country, ".xlsx")
 
  read_excel(file_path, sheet = 'ISI', range = 'A2:W10',
             col_names = c('product', 2000:2021)) |>
  mutate(country_code = country) |> 
  slice(7:8) |> 
  mutate(product = replace(product,
                           1:2, c('steel-primary', 'steel-secondary'))) |> 
  pivot_longer(cols ='2000':'2021', names_to = 'year', values_to = 'production_kt_idees')
}

test_list_isi_allyear <- lapply(country_codes$country_iso_code, read_data_isi_allyear)
isi_production_allyear <- bind_rows(test_list_isi_allyear, .id = "id") |> select(-id, year)

isi_production_allyear |> 
  pivot_wider(names_from = product, values_from = production_kt_idees) |>
  clean_names() |> 
  mutate(steel_total = rowSums(cbind(steel_primary, steel_secondary), na.rm = TRUE)) |> 
  pivot_longer(cols = steel_primary:steel_total, names_to = 'product_id') |> 
  mutate(mt = value/1000, .keep = 'unused') |> 
  filter(year %in% c(2010:2021))->  isi_production_allyear_plot
```
  
```{r, fig.width=10,fig.height=12}

  isi_production_allyear_plot |> 
  filter(product_id == 'steel_total') |> 
  ggplot(aes(year, mt))+
  geom_line(aes(group=1), linewidth= .8, color = 'red')+
  facet_wrap(~country_code, ncol = 5, scales = 'free')+
  theme_bw()+ 
  labs(title = 'Total Steel Historical Production in European Countries - JRC-IDEES',
       x = 'Year',
       y = 'Million Ton')+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) -> p50
  p50
  ggsave(here(output_dir, 'steel_production_total_allyears.pdf'), width = 14, height = 13, units = 'in', dpi = 900)
  
  ggsave(here(output_dir, 'steel_production_total_allyears.svg'), width = 16, height = 9, units = 'in', dpi = 900)
  
  isi_production_allyear_plot |> 
  filter(product_id == 'steel_primary') |> 
  ggplot(aes(year, mt))+
  geom_line(aes(group=1), linewidth= .8, color = 'orange')+
  facet_wrap(~country_code, ncol = 3, scales = 'free')+
  theme_bw()+
  labs(title = 'Primary Steel historical production in European Countries') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) -> p50
  p50
  ggsave(here(output_dir, 'steel_production_primary_allyears.pdf'), width = 14, height = 13, units = 'in', dpi = 900)
  
  ggsave(here(output_dir, 'steel_production_primary_allyears.svg'), width = 14, height = 13, units = 'in', dpi = 900)
  
  isi_production_allyear_plot |> 
  filter(product_id == 'steel_secondary') |> 
  ggplot(aes(year, mt))+
  geom_line(aes(group=1), linewidth= .8, color = 'blue')+
  facet_wrap(~country_code, ncol = 3, scales = 'free')+
  theme_bw()+
  labs(title = 'Secondary Steel historical production in European Countries') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) -> p50
  p50
  ggsave(here(output_dir, 'steel_production_secondary_allyears.pdf'), width = 14, height = 13, units = 'in', dpi = 900)
  
    ggsave(here(output_dir, 'steel_production_secondary_allyears.svg'), width = 14, height = 13, units = 'in', dpi = 900)

```

Production AIDRES

```{r}

 aidres_production_country_2018 |> 
        filter(sector_id == 'steel') |> 
        select(-sector_id, -production_route_name) |> 
    mutate(mt = round(production_kt_y/1000, 3), .keep = 'unused') |> 

ggplot(aes(mt, fct_reorder(country_code, mt, .fun = sum)))+
        geom_col(aes(fill = product_id), width = .6)+
        theme_bw()+
        labs(title = paste0('Steel sector production volume comparison for year 2018'),
             y = 'Country',
             x = 'Million Ton') + 
theme(legend.position = "bottom") +
theme(legend.title = element_blank())+
scale_fill_brewer(palette = "Set2") -> p50
  p50
  
ggsave(here(output_dir, 'steel_production_aidres_2018.pdf'), width = 14, height = 13, units = 'in', dpi = 900)
  
ggsave(here(output_dir, 'steel_production_aidres_2018.svg'), width = 14, height = 8, units = 'in', dpi = 900)
```

Production AIDRES 2.0

```{r}

 aidres_2_0_production_country_2018 |> 
        filter(sector_id == 'steel') |> 
        select(-sector_id, -production_route_name) |> 
    mutate(mt = round(production_kt_y/1000, 3), .keep = 'unused') |> 

ggplot(aes(mt, fct_reorder(country_code, mt, .fun = sum)))+
        geom_col(aes(fill = product_id), width = .6)+
        theme_bw()+
        labs(title = paste0('Steel sector production volume comparison for year 2018 2.0'),
             y = 'Country',
             x = 'Million Ton') + 
theme(legend.position = "bottom") +
theme(legend.title = element_blank())+
scale_fill_brewer(palette = "Set2") -> p50
  p50
  
ggsave(here(output_dir, 'steel_production_aidres_2018_2_0.pdf'), width = 14, height = 13, units = 'in', dpi = 900)
  
ggsave(here(output_dir, 'steel_production_aidres_2018_2_0.svg'), width = 14, height = 8, units = 'in', dpi = 900)
```


### Comparison of Steel Production between AIDRES/AIDRES 2.0 and JRC-IDEES

We compare AIDRES/AIDRES 2.0 steel production with JRC-IDEES for the year 2018.

#### Plot

```{r, fig.width=10, fig.height=7}
#| warning: false

isi_production_allyear_plot |> 
  mutate(source = 'idees') |> 
  mutate(mt = round(mt,3)) -> isi_production_allyear_plot

  aidres_production_country_2018 |> 
        filter(sector_id == 'steel') |> 
        select(-sector_id, -production_route_name) |> 
    mutate(mt = round(production_kt_y/1000, 3), .keep = 'unused') |> 
    pivot_wider(names_from = product_id, values_from = mt) |>
    clean_names() |> 
    mutate(steel_total = rowSums(cbind(steel_primary, steel_secondary), na.rm = TRUE)) |> 
    pivot_longer(cols = steel_primary:steel_total, names_to = 'product_id', values_to = 'mt') |> 
    mutate(source = 'aidres')  -> aidres_production_country_plot
  
  
    bind_rows(isi_production_allyear_plot |> filter(year == 2018) |> select(-year), 
            aidres_production_country_plot) |> 
    ggplot(aes(mt, fct_reorder(country_code, mt, .fun = sum)))+
        geom_line()+
        geom_point(aes(colour = source), size = 1.5)+
        facet_wrap(~product_id, ncol = 3)+
        theme_bw()+
        labs(title = paste0('Steel sector production volume comparison for year 2018'),
             y = 'Country',
             x = 'Million Ton') +
theme(legend.position = "bottom") +
theme(legend.title = element_blank()) -> p60

p60
    
ggsave(here(output_dir, paste0('steel_production_compare', 2018, '.pdf')), width = 10, height = 8, units = 'in', dpi = 900)

ggsave(here(output_dir, paste0('steel_production_compare', 2018, '.svg')), width = 14, height = 8, units = 'in', dpi = 900)

```

```{r, fig.width=10, fig.height=7}
#| warning: false

isi_production_allyear_plot |> 
  mutate(source = 'idees') |> 
  mutate(mt = round(mt,3)) -> isi_production_allyear_plot

  aidres_2_0_production_country_2018 |> 
        filter(sector_id == 'steel') |> 
        select(-sector_id, -production_route_name) |> 
    mutate(mt = round(production_kt_y/1000, 3), .keep = 'unused') |> 
    pivot_wider(names_from = product_id, values_from = mt) |>
    clean_names() |> 
    mutate(steel_total = rowSums(cbind(steel_primary, steel_secondary), na.rm = TRUE)) |> 
    pivot_longer(cols = steel_primary:steel_total, names_to = 'product_id', values_to = 'mt') |> 
    mutate(source = 'aidres')  -> aidres_2_0_production_country_plot
  
  
    bind_rows(isi_production_allyear_plot |> filter(year == 2018) |> select(-year), 
            aidres_2_0_production_country_plot) |> 
    ggplot(aes(mt, fct_reorder(country_code, mt, .fun = sum)))+
        geom_line()+
        geom_point(aes(colour = source), size = 1.5)+
        facet_wrap(~product_id, ncol = 3)+
        theme_bw()+
        labs(title = paste0('Steel sector production volume comparison for year 2018 2.0'),
             y = 'Country',
             x = 'Million Ton') +
theme(legend.position = "bottom") +
theme(legend.title = element_blank()) -> p60

p60
    
ggsave(here(output_dir, paste0('steel_production_compare', 2018, '.pdf')), width = 10, height = 8, units = 'in', dpi = 900)

ggsave(here(output_dir, paste0('steel_production_compare', 2018, '.svg')), width = 14, height = 8, units = 'in', dpi = 900)

```

#### Diff file

```{r}
bind_rows(isi_production_allyear_plot |> filter(year == 2018) |> 
            select(-year), aidres_production_country_plot) |> 
  mutate(year = 2018) |> 
  pivot_wider(values_from = mt, names_from = source) -> prod_comp_2018

write_xlsx(list('year_2018' = prod_comp_2018),
           here(output_dir, 'diff_production_volume_steel_2018.xlsx'))

```

```{r}
bind_rows(isi_production_allyear_plot |> filter(year == 2018) |> 
            select(-year), aidres_2_0_production_country_plot) |> 
  mutate(year = 2018) |> 
  pivot_wider(values_from = mt, names_from = source) -> prod_comp_2018_2_0

write_xlsx(list('year_2018' = prod_comp_2018_2_0),
           here(output_dir, 'diff_production_volume_steel_2018_2_0.xlsx'))

```

### Finishing and rolling energy

As AIDRES/AIDRES 2.0 does not include the energy consumption for refining, rolling, and finishing, these information are taken from JRC-IDEES database.

### Rolling energy by fuel

#### For primary steel

```{r}
#| warning: false

read_data_isi_r_fec_pri <- function(country){
  
  file_path <- paste0(idees_data, '/JRC-IDEES-2021_Industry_', country, ".xlsx")

  read_excel(file_path, sheet = 'ISI_fec', range = 'A2:W55',
             col_names = c('product', 2000:2021)) |> 
    select(1,20) |> 
    mutate(country_code = country) |> 
    slice(29, 30, 31, 32, 33) |> 
    mutate(fuel = case_when(product == 'LPG' ~ 'oil_petro_products',
                          product == 'Diesel oil and liquid biofuels'  ~ 'oil_petro_products',
                          product == 'Natural gas and biogas' ~ 'natural_gas',
                          product == 'Fuel oil' ~ 'oil_petro_products',
                          product == 'Steel: Furnaces, refining and rolling - Electric' ~ 'electricity')) |> 
    mutate(product = 'steel-primary') |>
    pivot_longer(cols ='2018', names_to = 'year', values_to = 'ktoe') |>
    mutate(rolling_gj_idees = ktoe*1000*41.87,
           ktoe = NULL) |> 
    group_by(country_code, product, fuel) |> 
    summarise(rolling_gj_idees=sum(rolling_gj_idees), .groups = 'drop') |> 
    mutate(country_share_roll = round(rolling_gj_idees/ sum(rolling_gj_idees), 3))
}

test_list_isi_r_fec_pri <- lapply(country_codes$country_iso_code, read_data_isi_r_fec_pri)
isi_r_fec_pri <- bind_rows(test_list_isi_r_fec_pri, .id = "id") |> select(-id)

```

#### For secondary steel

```{r}
#| warning: false

read_data_isi_r_fec_sec <- function(country){
  
file_path <- paste0(idees_data, '/JRC-IDEES-2021_Industry_', country, ".xlsx")
  
  read_excel(file_path, sheet = 'ISI_fec', range = 'A2:W100',
             col_names = c('product', 2000:2021)) |> 
    select(1,20) |> 
    mutate(country_code = country) |> 
    slice(73,74,75, 76, 77) |> 
    mutate(fuel = case_when(product == 'LPG' ~ 'oil_petro_products',
                          product == 'Diesel oil and liquid biofuels'  ~ 'oil_petro_products',
                          product == 'Natural gas and biogas' ~ 'natural_gas',
                          product == 'Fuel oil' ~ 'oil_petro_products',
                          product == 'Steel: Furnaces, refining and rolling - Electric' ~ 'electricity')) |> 
    mutate(product = 'steel-secondary') |>
    pivot_longer(cols ='2018', names_to = 'year', values_to = 'ktoe') |>
    mutate(rolling_gj_idees = ktoe*1000*41.87,
           ktoe = NULL) |> 
    group_by(country_code, product, fuel) |> 
    summarise(rolling_gj_idees=sum(rolling_gj_idees), .groups = 'drop') |> 
    mutate(country_share_roll = round(rolling_gj_idees/ sum(rolling_gj_idees), 3))
}

test_list_isi_r_fec_sec <- lapply(country_codes$country_iso_code, read_data_isi_r_fec_sec)
isi_r_fec_sec <- bind_rows(test_list_isi_r_fec_sec, .id = "id") |> select(-id)
```

#### Total rolling energy

```{r}
#| warning: false

isi_r_tot<- rbind(isi_r_fec_sec, isi_r_fec_pri) |> ungroup()

isi_tot_rolling <- isi_r_tot |> 
  group_by(country_code, product) |> 
  summarise(total = sum(rolling_gj_idees)) |> ungroup() |>
  pivot_longer(cols = total, names_to = 'fuel', values_to = 'rolling_gj_idees') |> 
  mutate(country_share_roll =1) |>  
  rbind(isi_r_tot)

isi_tot_rolling
```

### Finishing energy by fuel

#### For primary steel

```{r}
#| warning: false

read_data_isi_f_fec_pri <- function(country){
  
  file_path <- paste0(idees_data, '/JRC-IDEES-2021_Industry_', country, ".xlsx")
  
  read_excel(file_path, sheet = 'ISI_fec', range = 'A2:W100',
             col_names = c('product', 2000:2021)) |> 
    select(1,20) |> 
    mutate(country_code = country) |> 
    slice(36:38,40:48,50) |> 
    mutate(fuel = case_when(product == 'LPG' ~ 'oil_petro_products',
                          product == 'Diesel oil and liquid biofuels'  ~ 'oil_petro_products',
                          product == 'Natural gas and biogas' ~ 'natural_gas',
                          product == 'Solids' ~ 'solid_fossil_fuels',
                          product == 'Refinery gas' ~ 'oil_petro_products',
                          product == 'Fuel oil' ~ 'oil_petro_products',
                          product == 'Other liquids' ~ 'oil_petro_products',
                          product == 'Derived gases' ~ 'solid_fossil_fuels',
                          product == 'Biomass and waste' ~ 'renew_bio',
                          product == 'Steel: Product finishing - Electric' ~ 'electricity')) |> 
    mutate(product = 'steel-primary') |>
    pivot_longer(cols ='2018', names_to = 'year', values_to = 'ktoe') |>
    mutate(finish_gj_idees = ktoe*1000*41.87,
           ktoe = NULL) |> 
    group_by(country_code, product, fuel) |> 
    summarise(finish_gj_idees=sum(finish_gj_idees), .groups = 'drop') |> 
    mutate(country_share_finish = round(finish_gj_idees/ sum(finish_gj_idees), 3))
}

test_list_isi_f_fec_pri <- lapply(country_codes$country_iso_code, read_data_isi_f_fec_pri)
isi_f_fec_pri <- bind_rows(test_list_isi_f_fec_pri, .id = "id") |> select(-id)
```

#### For secondary steel

```{r}
#| warning: false

read_data_isi_f_fec_sec <- function(country){
  
  file_path <- paste0(idees_data, '/JRC-IDEES-2021_Industry_', country, ".xlsx")
  
  read_excel(file_path, sheet = 'ISI_fec', range = 'A2:W100',
             col_names = c('product', 2000:2021)) |> 
    select(1,20) |> 
    mutate(country_code = country) |> 
    slice(80:82,84:92,94) |> 
    mutate(fuel = case_when(product == 'LPG' ~ 'oil_petro_products',
                            product == 'Diesel oil and liquid biofuels'  ~ 'oil_petro_products',
                            product == 'Natural gas and biogas' ~ 'natural_gas',
                            product == 'Solids' ~ 'solid_fossil_fuels',
                            product == 'Refinery gas' ~ 'oil_petro_products',
                            product == 'Fuel oil' ~ 'oil_petro_products',
                            product == 'Other liquids' ~ 'oil_petro_products',
                            product == 'Derived gases' ~ 'solid_fossil_fuels',
                            product == 'Biomass and waste' ~ 'renew_bio',
                            product == 'Steel: Product finishing - Electric' ~ 'electricity')) |> 
    mutate(product = 'steel-secondary') |>
    pivot_longer(cols ='2018', names_to = 'year', values_to = 'ktoe') |>
    mutate(finish_gj_idees = ktoe*1000*41.87,
           ktoe = NULL) |> 
    group_by(country_code, product, fuel) |>
    summarise(finish_gj_idees=sum(finish_gj_idees), .groups = 'drop') |> 
    mutate(country_share_finish = round(finish_gj_idees/ sum(finish_gj_idees), 3))
}

test_list_isi_f_fec_sec <- lapply(country_codes$country_iso_code, read_data_isi_f_fec_sec)
isi_f_fec_sec <- bind_rows(test_list_isi_f_fec_sec, .id = "id") |> select(-id)
```

#### Total finishing energy

```{r}
#| warning: false

isi_f_tot<- rbind(isi_f_fec_sec, isi_f_fec_pri) |> ungroup()
  
isi_tot_finish <- isi_f_tot |> 
  group_by(country_code, product) |> 
  summarise(total = sum(finish_gj_idees)) |> ungroup() |>
  pivot_longer(cols = total, names_to = 'fuel', values_to = 'finish_gj_idees') |> 
  mutate(country_share_finish =1) |>  
  rbind(isi_f_tot)

isi_tot_finish
```

### Total rolling and finishing energy

```{r}

isi_tot_roll_finish <- left_join(isi_tot_finish, isi_tot_rolling, by = join_by(country_code, product, fuel))

isi_tot_roll_finish
```

### Calculate total rolling and finishing energy for AIDRES/AIDRES 2.0 Steel production

To calculate the total energy consumption for rolling and finishing of steel (for AIDRES production volume) we first calculate the specific energy consumption of finishing and rolling by fuel using 2018 steel production and energy consumption data of JRC-IDEES.

### Joining production volume, rolling, and finishing energy

Calculate fuel specific energy consumption factors for finishing and rolling

```{r}
steel_data_idees <- isi_production_2018 |> 
  left_join(isi_tot_roll_finish, by = join_by('product', 'country_code')) |> 
  mutate(gj_per_t_finish =  finish_gj_idees/ production_kt_idees/1000,
         gj_per_t_roll = rolling_gj_idees / production_kt_idees/1000)

steel_data_idees
```

Joining with AIDRES/AIDRES 2.0 production volume, calculating total finishing and rolling energy by fuel using AIDRES production volume and IDEES sec.

```{r}

aidres_finishing_rolling_pj <- aidres_production_country_2018 |> rename('product' = product_id,
                                         'production_kt_aidres' = production_kt_y) |> 
  filter(sector_id == 'steel') |> 
  select(-sector_id) |> 
  left_join(steel_data_idees, by = join_by('product', 'country_code')) |> 
  mutate(total_pj_finish_aidres = gj_per_t_finish*production_kt_aidres*1000/10^6,
         total_pj_roll_aidres = gj_per_t_roll*production_kt_aidres*1000/10^6) |> 
  select(-production_kt_idees, -finish_gj_idees, -country_share_finish) |> 
  select(country_code, product, fuel, everything())

aidres_finishing_rolling_pj
```

```{r}

aidres_2_0_finishing_rolling_pj <- aidres_2_0_production_country_2018 |> rename('product' = product_id,
                                         'production_kt_aidres' = production_kt_y) |> 
  filter(sector_id == 'steel') |> 
  select(-sector_id) |> 
  left_join(steel_data_idees, by = join_by('product', 'country_code')) |> 
  mutate(total_pj_finish_aidres = gj_per_t_finish*production_kt_aidres*1000/10^6,
         total_pj_roll_aidres = gj_per_t_roll*production_kt_aidres*1000/10^6) |> 
  select(-production_kt_idees, -finish_gj_idees, -country_share_finish) |> 
  select(country_code, product, fuel, everything())

aidres_2_0_finishing_rolling_pj
```

### Comparing Energy consumption of AIDRES/AIDRES 2.0 (with finishing and rolling from IDEES) with EUROSTAT

Summarizing finishing and rolling energy by country and fuel

```{r}
aidres_finishing_country_pj <- aidres_finishing_rolling_pj |> 
  group_by(country_code, fuel) |> 
  summarise(finishing_pj = sum(total_pj_finish_aidres),
            rolling_pj = sum(total_pj_roll_aidres), .groups = 'drop')

aidres_finishing_country_pj
```

```{r}
aidres_2_0_finishing_country_pj <- aidres_2_0_finishing_rolling_pj |> 
  group_by(country_code, fuel) |> 
  summarise(finishing_pj = sum(total_pj_finish_aidres),
            rolling_pj = sum(total_pj_roll_aidres), .groups = 'drop')

aidres_2_0_finishing_country_pj
```

Add finishing and rolling data with EUROSTAT and AIDRES/AIDRES 2.0 data

```{r}

data_all_finishing_rolling <- steel_eurostat_aidres |> 
  pivot_wider(names_from = source, values_from = pj) |> 
  left_join(aidres_finishing_country_pj, by = join_by(country_code,  fuel)) |> 
  mutate(aidres = rowSums(cbind(aidres, finishing_pj, rolling_pj), na.rm = TRUE)) |> 
  select(-finishing_pj, -rolling_pj) |> 
  pivot_longer(cols = eurostat:aidres, names_to = 'source', values_to = 'pj')

data_all_finishing_rolling
```

```{r}

data_all_finishing_rolling_2_0 <- steel_eurostat_aidres_2_0 |> 
  pivot_wider(names_from = source, values_from = pj) |> 
  left_join(aidres_2_0_finishing_country_pj, by = join_by(country_code,  fuel)) |> 
  mutate(aidres = rowSums(cbind(aidres, finishing_pj, rolling_pj), na.rm = TRUE)) |> 
  select(-finishing_pj, -rolling_pj) |> 
  pivot_longer(cols = eurostat:aidres, names_to = 'source', values_to = 'pj')

data_all_finishing_rolling_2_0
```

### Plotting and diff files

#### Plot

plotting with fuel categories

```{r,   fig.width=10, fig.height=12}
#|warning: false

data_all_finishing_rolling |> 
  ggplot(aes(pj, fct_reorder(country_code, pj, .fun = sum))) +
  geom_line()+
  geom_point(aes(colour = source), size = 1.5) +
  facet_wrap(~fuel, ncol = 5, scales = "free") +
  theme_bw() +
  labs(title = 'Fuel wise total energy consumption in the steel sector in AIDRES and EUROSTAT for the year 2018',
       subtitle = 'Rolling and finishing energy is included in AIDRES database from JRC-IDEES',
       y = 'country',
       x = 'PJ') +
theme(legend.title = element_blank())-> p4
  p4
  ggsave(here(output_dir, 'steel_fec_compare_fuel_fr.pdf'), width = 16, height = 8, units = 'in', dpi = 900)
  
    ggsave(here(output_dir, 'steel_fec_compare_fuel_fr.svg'), width = 12, height = 7, units = 'in', dpi = 900)
```

```{r,   fig.width=10, fig.height=12}
#|warning: false

data_all_finishing_rolling_2_0 |> 
  ggplot(aes(pj, fct_reorder(country_code, pj, .fun = sum))) +
  geom_line()+
  geom_point(aes(colour = source), size = 1.5) +
  facet_wrap(~fuel, ncol = 5, scales = "free") +
  theme_bw() +
  labs(title = 'Fuel wise total energy consumption in the steel sector in AIDRES 2.0 and EUROSTAT for the year 2018',
       subtitle = 'Rolling and finishing energy is included in AIDRES 2.0 database from JRC-IDEES',
       y = 'country',
       x = 'PJ') +
theme(legend.title = element_blank())-> p4
  p4
  ggsave(here(output_dir, 'steel_fec_compare_fuel_fr_2_0.pdf'), width = 16, height = 8, units = 'in', dpi = 900)
  
    ggsave(here(output_dir, 'steel_fec_compare_fuel_fr_2_0.svg'), width = 12, height = 7, units = 'in', dpi = 900)
```

plotting only total energy

```{r,  fig.width=10,fig.height=8}
#|warning: false

  data_all_finishing_rolling |> 
    filter(fuel == 'total') |> 
    ggplot(aes(pj, fct_reorder(country_code, pj, .fun = sum))) +
    geom_line()+
    geom_point(aes(colour = source), size = 1.5) +
    theme_bw() +
  labs(title = 'Total energy consumption in the steel sector in AIDRES and EUROSTAT',
       subtitle = 'Rolling and finishing energy is included in AIDRES database from JRC-IDEES for the year 2018',
       y = 'country',
       x = 'PJ') + 
    theme(legend.position = "bottom") +
    theme(legend.title = element_blank()) -> p5
  p5
   ggsave(here(output_dir, 'steel_fec_compare_tot_fr.pdf'), width = 10, height = 8, units = 'in', dpi = 900)
   
      ggsave(here(output_dir, 'steel_fec_compare_tot_fr.svg'), width = 12, height = 7, units = 'in', dpi = 900)
```

```{r,  fig.width=10,fig.height=8}
#|warning: false

  data_all_finishing_rolling_2_0 |> 
    filter(fuel == 'total') |> 
    ggplot(aes(pj, fct_reorder(country_code, pj, .fun = sum))) +
    geom_line()+
    geom_point(aes(colour = source), size = 1.5) +
    theme_bw() +
  labs(title = 'Total energy consumption in the steel sector in AIDRES 2.0 and EUROSTAT',
       subtitle = 'Rolling and finishing energy is included in AIDRES 2.0 database from JRC-IDEES for the year 2018',
       y = 'country',
       x = 'PJ') + 
    theme(legend.position = "bottom") +
    theme(legend.title = element_blank()) -> p5
  p5
   ggsave(here(output_dir, 'steel_fec_compare_tot_fr_2_0.pdf'), width = 10, height = 8, units = 'in', dpi = 900)
   
      ggsave(here(output_dir, 'steel_fec_compare_tot_fr_2_0.svg'), width = 12, height = 7, units = 'in', dpi = 900)
```

### Diff files

Fuel wise with finishing and rolling

```{r}
   data_all_finishing_rolling |>
   select( -country) |>
   pivot_wider(values_from = pj, names_from = source) |>
   write_xlsx(here(output_dir, 'diff_with_finishing_rolling_fuel.xlsx'))
```

```{r}
   data_all_finishing_rolling_2_0 |>
   select( -country) |>
   pivot_wider(values_from = pj, names_from = source) |>
   write_xlsx(here(output_dir, 'diff_with_finishing_rolling_fuel_2_0.xlsx'))
```

# Steel sector claibration w.r.t EUROSTAT data

We have obtained the rolling and finishing data from JRC-IDEES. We also calculate corresponding sec values. Now, we calibrate the AIDRES/AIDRES 2.0 final energy consumption w.r.t EUROSTAT values. Therefore, we subtract the total rolling and finishing energy from final energy  of EUROSTAT. This constitute the 'true' energy consumption for the process routes 'upto rolling and finishing'. Subsequently the sec factors are calculated for the process routes 'upto rolling and finishing'.

## Eurostat without finishing rolling

We subtract the finishing and rolling energy for AIDRES/AIDRES 2.0 steel production from total final energy consumption of EUROSTAT.

```{r}
   eurostat_without_finishing_rolling <- steel_eurostat_aidres |> 
    filter(sector == 'steel',
           source == 'eurostat') |> 
    select(-source) |> 
    mutate(pj = replace_na(pj, 0)) |> 
    left_join(aidres_finishing_country_pj, by = join_by('country_code', 'fuel')) |> 
    mutate(finishing_pj = replace_na(finishing_pj, 0),
           rolling_pj = replace_na(finishing_pj, 0)) |> 
    mutate(without_finishing_pj = pj - finishing_pj - rolling_pj) |> 
    select(-c(pj, finishing_pj, rolling_pj))

eurostat_without_finishing_rolling
```

```{r}
   eurostat_without_finishing_rolling_2_0 <- steel_eurostat_aidres_2_0 |> 
    filter(sector == 'steel',
           source == 'eurostat') |> 
    select(-source) |> 
    mutate(pj = replace_na(pj, 0)) |> 
    left_join(aidres_2_0_finishing_country_pj, by = join_by('country_code', 'fuel')) |> 
    mutate(finishing_pj = replace_na(finishing_pj, 0),
           rolling_pj = replace_na(finishing_pj, 0)) |> 
    mutate(without_finishing_pj = pj - finishing_pj - rolling_pj) |> 
    select(-c(pj, finishing_pj, rolling_pj))

eurostat_without_finishing_rolling_2_0
```

EUROSTAT doesn't have primary/ secondary breakup. This is obtained from IDEES data.

## Reading primary steel energy consumption from IDEES

```{r}
      read_data_isi_fec_pri <- function(country){
        
file_path <- paste0(idees_data, '/JRC-IDEES-2021_Industry_', country, ".xlsx")
        
    read_excel(file_path, sheet = 'ISI_fec', range = 'A2:W100',
               col_names = c('product', 2000:2021)) |> 
      select(1,20) |> 
      mutate(country_code = country) |> 
      slice(4, 34, 27) |> 
      rename(process = 1) |> 
      mutate(process = replace(process, 1:3, c('steel-primary-tot', 'steel-primary-finishing', 'steel-primary-rolling'))) |> 
      pivot_longer(cols ='2018', names_to = 'year', values_to = 'ktoe') |> 
      mutate(fec_gj_idees = ktoe*1000*41.87, ktoe = NULL) 
}
      
      test_list_isi_fec_pri <- lapply(country_codes$country_iso_code, read_data_isi_fec_pri)
      isi_fec_pri <- bind_rows(test_list_isi_fec_pri, .id = "id") |> select(-id, -year)
```

## Reading secondary steel energy consumption from IDEES

```{r}
      read_data_isi_fec_sec <- function(country){
        
file_path <- paste0(idees_data, '/JRC-IDEES-2021_Industry_', country, ".xlsx")
        
    read_excel(file_path, sheet = 'ISI_fec', range = 'A2:W100',
               col_names = c('product', 2000:2021)) |> 
      select(1,20) |> 
          mutate(country_code = country) |> 
          slice(53, 78, 71) |> 
          rename(process = 1) |> 
          mutate(process = replace(process,
                                   1:3, c('steel-secondary-tot', 'steel-secondary-finishing', 'steel-secondary-rolling'))) |> 
          pivot_longer(cols ='2018', names_to = 'year', values_to = 'ktoe') |> 
          mutate(fec_gj_idees = ktoe*1000*41.87,
                 ktoe = NULL)
      }
      
      test_list_isi_fec_sec <- lapply(country_codes$country_iso_code, read_data_isi_fec_sec)
      isi_fec_sec <- bind_rows(test_list_isi_fec_sec, .id = "id") |> select(-id, -year)
```

## Separating out finishing and rolling energy from total total final energy consumption in IDEES

```{r}

isi_fec_sec_fr <- isi_fec_sec |> 
  pivot_wider(names_from = process, values_from = fec_gj_idees) |> 
        clean_names() |> 
        mutate(fec_sec_net = steel_secondary_tot -  steel_secondary_finishing - steel_secondary_rolling) |> 
        select(-c(steel_secondary_tot, steel_secondary_finishing, steel_secondary_rolling))

      isi_fec_sec_fr

isi_fec_pri_fr <- isi_fec_pri |> 
        pivot_wider(names_from = process, values_from = fec_gj_idees) |> 
        clean_names() |> 
        mutate(fec_pri_net = steel_primary_tot - steel_primary_finishing - steel_primary_rolling) |> 
        select(-c(steel_primary_tot, steel_primary_finishing, steel_primary_rolling))
      
      isi_fec_pri_fr
```

## Ratio of secondary and primary energy consumption without finishing and rolling

```{r}

left_join(isi_fec_pri_fr, isi_fec_sec_fr, by = join_by('country_code')) |> 
        rename('steel_primary' = fec_pri_net,
               'steel_secondary' = fec_sec_net) |> 
        pivot_longer(cols = -country_code, names_to = 'prod_route', values_to = 'gj') |> 
        mutate(tot_pj = gj/10^6, .keep = 'unused') |> 
        group_by(country_code) |> 
        mutate(prod_route_share = tot_pj/ sum(tot_pj)) |> 
        select(-tot_pj) |> 
        mutate(prod_route = str_remove(prod_route, '_tot'))-> steel_idees_fec_share_no_finish_roll

steel_idees_fec_share_no_finish_roll
```

Applying primary, secondary energy (without rolling and finishing) to the EUROSTAT FEC (without rolling and finishing) to obtain primary, secondary breakup of energy consumption (without rolling and finishing)

```{r}

eurostat_without_finishing_rolling |> filter(sector == 'steel',
                                             fuel == 'total') |> 
  select(-sector, -fuel) -> steel_eurostat_fec_no_finish
 
left_join(steel_idees_fec_share_no_finish_roll, steel_eurostat_fec_no_finish, by = join_by('country_code')) |> 
        select(country_code, country, prod_route, prod_route_share, without_finishing_pj) |> 
        mutate(pj_route = without_finishing_pj*prod_route_share, .keep = 'unused') -> steel_eurostat_fec_prod_route
steel_eurostat_fec_prod_route
```

```{r}

eurostat_without_finishing_rolling_2_0 |> filter(sector == 'steel',
                                             fuel == 'total') |> 
  select(-sector, -fuel) -> steel_eurostat_fec_no_finish_2_0
 
left_join(steel_idees_fec_share_no_finish_roll, steel_eurostat_fec_no_finish_2_0, by = join_by('country_code')) |> 
        select(country_code, country, prod_route, prod_route_share, without_finishing_pj) |> 
        mutate(pj_route = without_finishing_pj*prod_route_share, .keep = 'unused') -> steel_eurostat_fec_prod_route_2_0
steel_eurostat_fec_prod_route_2_0
```

#### New SECs for production routes no finishing and roling

Applying EUROSTAT fuel share of primary and secondary to the net energy

```{r}

steel_eurostat_aidres |> 
  filter(sector == 'steel',
         source == 'eurostat') |> 
  select(- sector, -source) |> 
  pivot_wider(names_from = fuel, values_from = pj) |> 
  mutate(solid_fossil_fuels  = solid_fossil_fuels / total,
         electricity  = electricity / total,
         natural_gas = natural_gas / total,
         oil_petro_products = oil_petro_products / total,
         renew_bio = renew_bio / total,
         non_renew_waste = non_renew_waste / total,
         hydrogen  = hydrogen / total,
         methanol = methanol / total,
         heat = heat/ total) |> 
  select(-total) |> 
  pivot_longer(cols = electricity:methanol, names_to = 'fuel', values_to = 'fuel_share') |> 

left_join(steel_eurostat_fec_prod_route, by=join_by(country_code, country), relationship = 'many-to-many') |> 
  select(country_code, country, prod_route, everything()) |> 
  mutate(fuel_route_pj = round(pj_route*fuel_share,3)) -> steel_fuel_route_pj
steel_fuel_route_pj
```

```{r}

steel_eurostat_aidres_2_0 |> 
  filter(sector == 'steel',
         source == 'eurostat') |> 
  select(- sector, -source) |> 
  pivot_wider(names_from = fuel, values_from = pj) |> 
  mutate(solid_fossil_fuels  = solid_fossil_fuels / total,
         electricity  = electricity / total,
         natural_gas = natural_gas / total,
         oil_petro_products = oil_petro_products / total,
         renew_bio = renew_bio / total,
         non_renew_waste = non_renew_waste / total,
         hydrogen  = hydrogen / total,
         methanol = methanol / total,
         heat = heat/ total) |> 
  select(-total) |> 
  pivot_longer(cols = electricity:methanol, names_to = 'fuel', values_to = 'fuel_share') |> 

left_join(steel_eurostat_fec_prod_route_2_0, by=join_by(country_code, country), relationship = 'many-to-many') |> 
  select(country_code, country, prod_route, everything()) |> 
  mutate(fuel_route_pj = round(pj_route*fuel_share,3)) -> steel_fuel_route_pj_2_0
steel_fuel_route_pj_2_0
```

#### Specific energy consumption factor upto finishing and rolling

```{r}

isi_aidres_production_comp |> 
  filter(product_id!= 'steel_total') -> steel_prod_aidres
  
left_join(steel_fuel_route_pj, steel_prod_aidres, by = join_by('country_code' == 'country_code', 'prod_route' == 'product_id')) |>
        select(-source) |> 
      mutate(sec_gj_t = fuel_route_pj*10^6/(kt*1000)) |> 
       select(-c(fuel_share, pj_route, fuel_route_pj, kt)) -> steel_aidres_fec_no_finish_roll
      
      steel_aidres_fec_no_finish_roll
```

```{r}

isi_aidres_2_0_production_comp |> 
  filter(product_id!= 'steel_total') -> steel_prod_aidres_2_0
  
left_join(steel_fuel_route_pj_2_0, steel_prod_aidres_2_0, by = join_by('country_code' == 'country_code', 'prod_route' == 'product_id')) |>
        select(-source) |> 
      mutate(sec_gj_t = fuel_route_pj*10^6/(kt*1000)) |> 
       select(-c(fuel_share, pj_route, fuel_route_pj, kt)) -> steel_aidres_2_0_fec_no_finish_roll
      
      steel_aidres_2_0_fec_no_finish_roll
```

#### Secs for finishing and rolling

```{r}

aidres_finishing_rolling_pj |> filter(fuel != 'total') |> 
  select(country_code, product, production_route_name, fuel, gj_per_t_finish, gj_per_t_roll) |> 
  rename('sec_finish_gj_t' = gj_per_t_finish,
         'sec_roll_gj_t' = gj_per_t_roll) |> 
  mutate(product=replace(product, product=='steel-primary', 'steel_primary')) |> 
  mutate(product= replace(product, product == 'steel-secondary', 'steel_secondary')) -> sec_rolling_finish

sec_rolling_finish |> 
  select(-sec_finish_gj_t) |> 
  mutate(production_route_name = paste0('rolling-', production_route_name)) |> 
  pivot_wider(names_from = fuel, values_from = sec_roll_gj_t) -> sec_rolling_steel

sec_rolling_steel

sec_rolling_finish |> 
  select(-sec_roll_gj_t) |> 
  mutate(production_route_name = paste0('finishing-', production_route_name)) |> 
  pivot_wider(names_from = fuel, values_from = sec_finish_gj_t) -> sec_finishing_steel

sec_finishing_steel
```

```{r}

aidres_2_0_finishing_rolling_pj |> filter(fuel != 'total') |> 
  select(country_code, product, production_route_name, fuel, gj_per_t_finish, gj_per_t_roll) |> 
  rename('sec_finish_gj_t' = gj_per_t_finish,
         'sec_roll_gj_t' = gj_per_t_roll) |> 
  mutate(product=replace(product, product=='steel-primary', 'steel_primary')) |> 
  mutate(product= replace(product, product == 'steel-secondary', 'steel_secondary')) -> sec_rolling_finish_2_0

sec_rolling_finish_2_0 |> 
  select(-sec_finish_gj_t) |> 
  mutate(production_route_name = paste0('rolling-', production_route_name)) |> 
  pivot_wider(names_from = fuel, values_from = sec_roll_gj_t) -> sec_rolling_steel_2_0

sec_rolling_steel_2_0

sec_rolling_finish_2_0 |> 
  select(-sec_roll_gj_t) |> 
  mutate(production_route_name = paste0('finishing-', production_route_name)) |> 
  pivot_wider(names_from = fuel, values_from = sec_finish_gj_t) -> sec_finishing_steel_2_0

sec_finishing_steel_2_0
```

#### Adding finishing and rolling secs to the database

```{r}

aidres_finishing_rolling_pj |> filter(fuel != 'total') |> 
  select(country_code, product, production_route_name, fuel, gj_per_t_finish, gj_per_t_roll) |> 
  rename('sec_finish_gj_t' = gj_per_t_finish,
         'sec_roll_gj_t' = gj_per_t_roll) |> 
  mutate(product=replace(product, product=='steel-primary', 'steel_primary')) |> 
  mutate(product= replace(product, product == 'steel-secondary', 'steel_secondary')) -> aidres_finishing_rolling_pj_
  
steel_aidres_fec_no_finish_roll |> left_join(aidres_finishing_rolling_pj_, by = join_by('country_code' == 'country_code','prod_route' == 'product','fuel' == 'fuel','production_route_name' == 'production_route_name')) |> 
  rename('sec_norf_gj_t' = sec_gj_t) |> 
  select(country_code, country, everything()) |> 
  arrange(country_code , prod_route ) |> 
  filter(!is.na(production_route_name))-> steel_sec
  
  steel_sec
```

```{r}

aidres_2_0_finishing_rolling_pj |> filter(fuel != 'total') |> 
  select(country_code, product, production_route_name, fuel, gj_per_t_finish, gj_per_t_roll) |> 
  rename('sec_finish_gj_t' = gj_per_t_finish,
         'sec_roll_gj_t' = gj_per_t_roll) |> 
  mutate(product=replace(product, product=='steel-primary', 'steel_primary')) |> 
  mutate(product= replace(product, product == 'steel-secondary', 'steel_secondary')) -> aidres_2_0_finishing_rolling_pj_
  
steel_aidres_2_0_fec_no_finish_roll |> left_join(aidres_2_0_finishing_rolling_pj_, by = join_by('country_code' == 'country_code','prod_route' == 'product','fuel' == 'fuel','production_route_name' == 'production_route_name')) |> 
  rename('sec_norf_gj_t' = sec_gj_t) |> 
  select(country_code, country, everything()) |> 
  arrange(country_code , prod_route ) |> 
  filter(!is.na(production_route_name))-> steel_sec_2_0
  
  steel_sec_2_0
```

#### Creating final data frame for steel

```{r}
steel_sec |>
  mutate(unit = 'gj_t') |> 
  rename('finishing' = sec_finish_gj_t ,
         'rolling' = sec_roll_gj_t ,
         'upto_roll_fin' = sec_norf_gj_t,
         'prod_route' = production_route_name,
         'product' =  'prod_route') |> 
  pivot_longer(cols = c(finishing, upto_roll_fin,rolling) ,  names_to = 'process_route', values_to = 'value') |> 
  mutate(value = round(value, 3)) -> steel_sec1
  
steel_sec1 |> 
  filter(process_route == 'upto_roll_fin') |> 
  mutate(process_route = paste(prod_route, process_route, sep = '-')) |> 
   select(-prod_route) |> 
  pivot_wider(names_from = fuel, values_from = value) |> 
  select(country_code:product, process_route, unit, everything()) -> steel_sec_upto_fr

steel_sec_upto_fr
  
steel_sec1 |> 
  filter(process_route != 'upto_roll_fin') |> 
  select(-prod_route,) |> 
  pivot_wider(names_from = fuel, values_from = value) |> 
  select(country_code:product, process_route, unit, everything()) -> steel_sec_fr

steel_sec_fr

write_xlsx(list('sec_upto_rolling_finishing' = steel_sec_upto_fr,
                'sec_finishing_rolling' = steel_sec_fr),
          here(output_dir, 'steel_final_process_routes.xlsx') )
```

```{r}
steel_sec_2_0 |>
  mutate(unit = 'gj_t') |> 
  rename('finishing' = sec_finish_gj_t ,
         'rolling' = sec_roll_gj_t ,
         'upto_roll_fin' = sec_norf_gj_t,
         'prod_route' = production_route_name,
         'product' =  'prod_route') |> 
  pivot_longer(cols = c(finishing, upto_roll_fin,rolling) ,  names_to = 'process_route', values_to = 'value') |> 
  mutate(value = round(value, 3)) -> steel_sec1_2_0
  
steel_sec1_2_0|> 
  filter(process_route == 'upto_roll_fin') |> 
  mutate(process_route = paste(prod_route, process_route, sep = '-')) |> 
   select(-prod_route) |> 
  pivot_wider(names_from = fuel, values_from = value) |> 
  select(country_code:product, process_route, unit, everything()) -> steel_sec_upto_fr_2_0

steel_sec_upto_fr_2_0
  
steel_sec1_2_0 |> 
  filter(process_route != 'upto_roll_fin') |> 
  select(-prod_route,) |> 
  pivot_wider(names_from = fuel, values_from = value) |> 
  select(country_code:product, process_route, unit, everything()) -> steel_sec_fr_2_0

steel_sec_fr_2_0

write_xlsx(list('sec_upto_rolling_finishing' = steel_sec_upto_fr_2_0,
                'sec_finishing_rolling' = steel_sec_fr_2_0),
          here(output_dir, 'steel_final_process_routes.xlsx') )
```

### Writing production routes in a excel workbook

```{r}
#| eval: false
sec_upto_rf_new |> 
  write.xlsx(here(output_dir, 'steel_production_routes.xlsx'), sheet = 'sec_upto_rf_new')

sec_rolling_steel |> 
    write.xlsx(here(output_dir, 'steel_production_routes.xlsx'), sheet = 'sec_rolling', append=TRUE)

sec_finishing_steel |> 
    write.xlsx(here(output_dir, 'steel_production_routes.xlsx'), sheet = 'sec_finishing', append=TRUE)
```
